Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-04-20 paper.tex
Final corrections before sending to author-list

As mentioned the understanding of self-sustained oscillations on supported
catalysts is hampered by the fact that entirely different mechanisms than those
observed on single and polycrystalline samples may exist. For example it was
shown on a Pt(110) crystal how waves of surface reconstruction could propagate
on the surface ensuring a synchronization of the phenomena. Synchronization also have to
occur for oscillations on supported catalysts but here it is more likely to be
due to coupling of mass or heat transport. There have been a few examples where
such oscillations have been observed. Of these the most prominent work is by
Skoghlund\cite{Carlsson2006} where self-sustained oscillations were observed on
a supported Pt catalyst while simultaneously studying the catalyst by in-situ
IR spectrometry. Similar studies were recently undertaken using EXAFS
\cite{Singh2010} but here the oscillatory behavior was unstable and less
convincing.

In this work we demonstrate how a two-dimensional model system of Pt
nanoparticles on a SiO$_2$ support can be constructed that shows highly
repeatable self-sustained oscillations when initially started. The system
consist of mass selected Pt nanoparticles in the range from 2 to 10\,nm 
deposited in a microreactor with a surface area of 0.75\,cm$^2$
especially designed to achieve high sensitivity\cite{Henriksen2009}. This
method offers some advantages over the chemical synthesis methods, which also
offers the possibility of size selected nanoparticles, see e.g.
\cite{Tsang2008,Shao2011}. Particularly, surfactants or solvents
\cite{Biacchi2011}, which are required for
chemical synthesis methods, are avoided with gas-aggregation formed particles
while still being able to vary the coverage from sub-percent to full coverage.
It is demonstrated how the nanoparticles provides self-sustained
oscillations over extend of days and how various parameters can be mapped out
including the temperature dependence -- an effect that could
not be demonstrated for a continuous film of same area and similar conditions.
By combining these results with previous studies \cite{Carlsson2006,Singh2010}
and a recent detailed study of CO oxidation on Pd using STM
\cite{Hendriksen2010} a plausible explanation for the oscillatory behavior is
presented. The low coverage and extremely efficient Pt nanoparticles illustrate
some of the pit falls that may have to be considered when making in-situ
investigations of such a system since extremely low coverages may provide the
entire conversion.

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off CO-
oxidation measurement is performed (shown in \ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at the same rate as the
heating ramp. After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in \ref{fgr:initial_treatment} where both the symmetrical
light-off ramp and the subsequent increase in temperature to achieve sustained
oscillations is seen. Often the sample will show oscillations either directly
upon light-off or on the falling temperature ramp.

@@

Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm
particles with a geometrical coverage of 0.1\% of the reactor area gave the
most stable oscillations. No consistent dependency on oscillation frequency and
duty cycle was found that could be attributed to the size of the particles.

The oscillations on all samples are qualitatively similar but the oscillation
period varies from a few seconds (the time-constant of the reactor is
$\sim7\,$s) up to more than an hour. Once a sample start to oscillate it will
perform self-sustained oscillations for as long as the experiment is allowed to
run. We have not observed a single sample stop oscillating once the self-
sustained oscillations had started.

An example of a single very slow oscillation is seen in
\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ (notice the CO signal does not go to zero due to
the cracking pattern of CO$_2$ and CO background in the QMS) the reaction shows
a fast deactivation. After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and convert all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

Due to the long timescales between full conversion and low conversion plateaus
the oscillations cannot be attributed to the reactor setup itself which due to
its small size has timescales on the order of seconds. Furthermore, a series of
experiments was performed to exclude the oscillations as an instrumental
artifact. Firstly, experiments were performed on a larger volume reactor thus
increasing the residence time in the reactor. Secondly, the oscillations were
reproduced on a comparable but not identical setup. Finally, attempts at
reproducing the phenomenon on Pt-thin films of comparable coverage did not
produce any oscillations. All of these experiments suggest that the
oscillations is a property of the Pt nanoparticles.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in \ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and increases linearly with time until 1500
min of total oscillation time. Hereafter the oscillations become more irregular
as shown in \ref{fgr:extracts} and lower panel \ref{fgr:long_measurement}. At
experiment end the oscillation period is between 20 and 45 min.

@@

\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min are
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in \ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 days long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in \ref{fgr:extracts}. This phenomena is  consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more short-
period oscillations despite the trend of increasingly slower oscillation
period. A possible explanation of the phenomeonon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time, i.e. the samples maintains a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
then be expected.

@@

\subsection{Temperature dependence}
\begin{figure}
\includegraphics[width=12cm]{temperature_dependence.png}
\caption{Illustration of the strong temperature dependence. After about 10
  hours of steady oscillations at 205$^\circ$C the temperature is lowered by
  5$^\circ$C immediately decreasing the oscillation frequency. The temperature
  is hereafter increased by 2$^\circ$C resulting in an increase in oscillation
  frequency. At 205$^\circ$C the oscillation period was approximately 500\,s
  while the oscillation period period is approximately 5000\,s. at
  203$^\circ$C
  \label{fgr:temperature_dependence}
}
\end{figure}

An interesting feature in the data is the very large temperature dependence on
oscillation frequency, as shown in \ref{fgr:temperature_dependence}. By
changing the temperature 5$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependency where in some cases a provoked small change in temperature will
consistently turn on and off the oscillations by changing the temperature
$20^\circ$C, see \ref{fgr:temperature_dependence_supplemental}.
\begin{figure}
\centering
  \includegraphics[width=12cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence.}
  \label{fgr:temperature_dependence_supplemental}
\end{figure}
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

@@

Although oscillations have previously been observed on extended Pt thin films
\cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric
pressure no oscillations could be provoked on thin films of comparable
geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

\subsection{Duty cycle}
Even though the period of the oscillations increases with time, the total
integrated conversion rate during a complete cycle is almost constant. In
\ref{fgr:duty_cycles_supplemental}, the value of the mean value of CO and
CO$_2$ is plotted for every oscillation in the four-day long experiment. It is
evident that the ratio between CO and CO$_2$ is almost constant and thus the
average activity of the sample is almost unchanged despite the development in
the oscillation frequency.

@@

\subsection{Size dependence}
Nanoparticles of several different sizes were tested and all showed
oscillations. However, no change in duty cycle or frequency that could be
attributed to the size of the nanoparticles was found. This may be due to the
fact that even nominally identical samples also show a large variation in
oscillation frequency and duty cycle thus possibly shadowing a nanoparticle
size effect.

\subsection{Proposed mechanism}
In this study Pt nanoparticles were investigated which, compared to a single
crystal or a thin film, is a very rough surface. It is to be expected that the
particles will initially favor the reduced state. As the reaction runs the gas
composition towards the outlet of the reactor will become more oxidizing
increasing the oxidation rate of the particles (illustrated in
\ref{fgr:full_oscillation}). Gradually, as the particles towards the outlet
oxidizes the turn-over will increase and the general CO concentration will
decrease promoting more and more particles to oxidize which will gradually be
seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO
concentration decreases the light-off temperature will decrease hence
approaching the constant temperature of the sample resulting in a steep
increase in reactivity. It is important to note that the fraction of oxidized
particles needed to achieve full conversion is not known. We are currently
investigating new methods for gaining insight into the nature of the
nanoparticles in the reactor but the low coverage and the design of the reactor
makes this a challenging task. If not all particles are oxidized at light-off
they will now oxidize much faster due to the low CO-concentration. The further
increased rate will not be visible in the QMS since the reactor is already in
full conversion. As most of the particles are now oxidized they will become
more and more roughened and gradually return to the reduced state. Fewer and
fewer oxidized particles will be responsible for the activity increasing the
roughing rate of the remaining active particles. Just before the overall
activity drops only a fraction of the particles participates in the reaction
and the deactivation will happen very suddenly. Once these potentially few
particles eventually becomes sufficiently rough they will reduce and lose
activity hence completing the oscillation cycle. It should be noted that our
group has recently measured turnover frequencies of CO oxidation on Pt in
excess of $10^{6}$s$^{-1}$ per site corresponding to as little as $\sim$300 Pt
nanoparticles contributing to the reactivity just before the end of the high-
activity cycle.

\section{Conclusions}
Self-sustained CO oxidation oscillations at atmospheric pressure on Pt
nanoparticles ranging from from 3\,nm to 9\,nm have been measured with high-
quality mass-spectrometry. It was found that the samples increased the
oscillation period over time and showed a strong temperature dependence. A
model is proposed where the nanoparticles go through repetitive
oxidation/reduction cycles which change the change rate of CO oxidation on the
particles.

\section{Experimental Methods}
\subsection{The $\mu$-reactor platform}
The experiments were performed in Si-based $\mu$-reactors\cite{Henriksen2009}.
The reactor design makes sure that all gases exposed to the catalyst under
investigation is measured by the quadropole mass spectrometer (QMS) ensuring an
extremely high sensitivity of the system.

The experiments were performed in Si-based 20x16\,mm
$\mu$-reactors\cite{Henriksen2009}. The reactor consists of two inlets, a
mixing zone allowing for diffusional mixing of reactants, an outlet, a reactor
volume of $240\,$nL and a $5.4\,\mu$m wide, 3\,$\mu$m high and 1500\,$\mu$m
long capillary used for sniffing gases from the reactor volume. The reactor is
sealed by anodic bonding of a pyrex lid to the Si reactor and is able to
operate in a pressure range of 0--2.5\,bar. The reaction gases are supplied to
the two inlets by flow controllers capable of controlling the gas flow from 0--
10 mL/min. The capillary flow is fed into a quadropole mass-spectrometer (QMS)
for analysis while any surplus of gas from the inlets is passed directly
through to the outlet via a pressure controller allowing control of reactor
pressure. The design makes sure that all gases exposed to the catalyst under
investigation are measured by the QMS ensuring an extremely high sensitivity of
the system. \begin{figure} \includegraphics[width=8cm]{Ib-reactor.png}
\caption{Image of the 20x16\,mm microreactor showing the circular reactor area
and gas channel system} \label{fgr:reactor} \end{figure} Before any
measurements are performed the reactor is pumped by a turbopump to minimize
contaminants in the system. When an evacuated reactor is mounted the base
pressure of the mass spectrometer chamber is $\sim5\times10^{-9}\,$mbar and in
operation at $1\,$bar in the reactor volume the pressure is
$\sim5\times10^{-7}\,$mbar in the QMS chamber.

@@

\subsection{Pt nanoparticle deposition}
Pt nanoparticles were deposited in the reactor using a gas-aggregation source
(Mantis Deposition Ltd., Nanogen 50). Pt clusters were formed by gas-phase
condensation of Pt atoms produced by impinging argon ions on a 99.99\% Pt
target in a magnetron sputter source. After condensation the ionised fraction
(60-80\%) of the clusters were size-selected by a quadropole mass selection
filter according to their mass-to-charge ratio. Using this setup Pt
nanoparticles with diameters in the range of 2--16\,nm can be produced
\cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after
size-selection soft-landed in the reactor volume of the microreactor. The
coverage was kept at approximately 0.1\% geometric coverage determined by
measuring the current on the reactor during deposition. After deposition the
reactors were anodically cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid
sintering of the nanoparticles while bonding.

>>>>

The understanding of self-sustained oscillations on supported
catalysts is hampered by the fact that entirely different mechanisms than those
observed on single and polycrystalline samples may exist. For example it was
shown on a Pt(110) crystal how waves of surface reconstruction could propagate
on the surface ensuring a synchronization of the phenomena. Synchronization also have to
occur for oscillations on supported catalysts but here it is more likely to be
due to coupling of mass or heat transport. There have been a few examples where
such oscillations have been observed. Of these the most prominent work is by
Skoghlund\cite{Carlsson2006} where self-sustained oscillations were observed on
a supported Pt catalyst while simultaneously studying the catalyst by in-situ
IR spectrometry. Similar studies were recently undertaken using EXAFS
\cite{Singh2010} but here the oscillatory behavior was unstable and less
convincing.

In this work we demonstrate how a two-dimensional model system of Pt
nanoparticles on a SiO$_2$ support can be constructed that shows highly
repeatable self-sustained oscillations when initially started. The system
consists of mass selected Pt nanoparticles in the range from 2 to 10\,nm 
deposited in a microreactor with a surface area of 0.75\,cm$^2$
especially designed to achieve high sensitivity\cite{Henriksen2009}. This
method offers some advantages over the chemical synthesis methods, which also
offers the possibility of size selected nanoparticles, see e.g.
\cite{Tsang2008,Shao2011}. Particularly, surfactants or solvents
\cite{Biacchi2011}, which are required for
chemical synthesis methods, are avoided with gas-aggregation formed particles
while still being able to vary the coverage from sub-percent to full coverage.
It is demonstrated how the nanoparticles provides self-sustained
oscillations stable in course of several days and how various parameters can be mapped out
including the temperature dependence -- an effect that, to our knowledge, has
never been measured before and that could
not be demonstrated for a continuous film of same area and similar conditions.
By combining these results with previous studies \cite{Carlsson2006,Singh2010}
and a recent detailed study of CO oxidation on Pd using STM
\cite{Hendriksen2010} a plausible explanation for the oscillatory behavior is
presented. The low coverage and extremely efficient Pt nanoparticles illustrate
some of the pit falls that may have to be considered when making in-situ
investigations of such a system since extremely low coverages may provide the
entire conversion.

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off
CO-oxidation measurement is performed (shown in \ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At a CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of the reactor area of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at the same rate as the
heating ramp. After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in \ref{fgr:initial_treatment} where both the symmetrical
light-off ramp and the subsequent increase in temperature to achieve sustained
oscillations is seen. Often the sample will show oscillations either directly
upon light-off or on the falling temperature ramp.

@@

The oscillations on all samples are qualitatively similar but the oscillation
period varies from a few seconds (the time constant of the reactor is
$\sim7\,$s) up to more than an hour. Once a sample start to oscillate it will
perform self-sustained oscillations for as long as the experiment is allowed to
run. We have not observed a single sample stop oscillating once the self-sustained oscillations had started.

An example of a single very slow oscillation is seen in
\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ the reaction shows a fast deactivation. 
Notice the CO signal does not go to zero due to the cracking pattern of
CO$_2$ and CO background in the Quadrupole Mass Spectrometer (QMS).
After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and convert all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

Due to the long timescales between full conversion and low conversion plateaus
the oscillations cannot be attributed to the reactor setup itself which due to
its small size has timescales on the order of seconds. Furthermore, a series of
experiments were performed to exclude the oscillations as an instrumental
artifact. Firstly, experiments were performed on a larger volume reactor thus
increasing the residence time in the reactor. Secondly, the oscillations were
reproduced on a comparable but not identical setup. Finally, attempts at
reproducing the phenomenon on Pt thin films of comparable coverage did not
produce any oscillations. All of these experiments suggest that the
oscillations is a property of the Pt nanoparticles. However, 
although oscillations have previously been observed on extended Pt thin films
\cite{Singh2010} as well as single crystals\cite{Hendriksen2005} at atmospheric
pressure no oscillations could be provoked on thin films of comparable
geometrical coverage to the nanoparticles in our system at any CO/O$_2$ ratio.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in \ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and increases linearly with time until 1500\,min
of total oscillation time. Hereafter the oscillations become more irregular
as shown in \ref{fgr:long_measurement} and in the lower panel of \ref{fgr:extracts}. At
experiment end the oscillation period is between 20\,min and 45\,min.

@@

\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min is
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in \ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from a 4 day long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in \ref{fgr:extracts}. This phenomena is  consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more
short-period oscillations despite the trend of increasingly slower oscillation
period. A possible explanation of the phenomeonon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time, i.e. the samples maintain a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
be expected.

@@

\subsection{Temperature dependence}
\begin{figure}
\includegraphics[width=12cm]{temperature_dependence.png}
\caption{Illustration of the strong temperature dependence. After about 10
  hours of steady oscillations at 205\,$^\circ$C the temperature is lowered by
  5\,$^\circ$C immediately decreasing the oscillation frequency. The temperature
  is hereafter increased by 2\,$^\circ$C resulting in an increase in oscillation
  frequency. At 205\,$^\circ$C the oscillation period was approximately 500\,s
  while the oscillation period period is approximately 5000\,s. at
  203\,$^\circ$C
  \label{fgr:temperature_dependence}
}
\end{figure}

An interesting feature in the data is the very large temperature dependence on
oscillation frequency as shown in \ref{fgr:temperature_dependence}. By
changing the temperature 5\,$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependence where in some cases a provoked small change in temperature will
consistently turn on and off the oscillations, see \ref{fgr:temperature_dependence_supplemental}.
\begin{figure}
\centering
  \includegraphics[width=12cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence.
  here the oscillations are turned on and off by changing the temperature
  $20^\circ$C,}
  \label{fgr:temperature_dependence_supplemental}
\end{figure}
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

@@

@@

\subsection{Duty cycle}
Even though the period of the oscillations increases with time the total
integrated conversion rate during a complete cycle is almost constant. In
\ref{fgr:duty_cycles_supplemental} the mean value of CO and
CO$_2$ is plotted for every oscillation in the four-day long experiment. It is
evident that the ratio between CO and CO$_2$ is almost constant. The
average activity of the sample is thus almost unchanged despite the development in
the oscillation frequency.

@@

\subsection{Size dependence}
Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm
particles with a geometrical coverage of 0.1\% of the reactor area gave the
most stable oscillations. 
However, no change in duty cycle or frequency that could be
attributed to the size of the nanoparticles was found. This may be due to the
fact that even nominally identical samples also show a large variation in
oscillation frequency and duty cycle thus possibly shadowing a nanoparticle
size effect.

\subsection{Proposed mechanism}
In this study Pt nanoparticles were investigated which, compared to a single
crystal or a thin film, is a very rough surface. It is to be expected that the
particles will initially favor the reduced state. As the reaction runs the gas
composition towards the outlet of the reactor will become more oxidizing
increasing the oxidation rate of the particles (illustrated in
\ref{fgr:full_oscillation}). Gradually, as the particles towards the outlet
oxidizes the turn-over frequency will increase and the general CO concentration will
decrease promoting more and more particles to oxidize which will gradually be
seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO
concentration decreases the light-off temperature will decrease hence
approaching the constant temperature of the sample resulting in a steep
increase in reactivity. It is important to note that the fraction of oxidized
particles needed to achieve full conversion is not known. We are currently
investigating new methods for gaining insight into the nature of the
nanoparticles in the reactor but the low coverage and the design of the reactor
makes this a challenging task. If not all particles are oxidized at light-off
they will now oxidize much faster due to the low CO-concentration. The further
increased rate will not be visible in the QMS since the reactor is already in
full conversion. As most of the particles are now oxidized they will become
more and more roughened and gradually return to the reduced state. Fewer and
fewer oxidized particles will be responsible for the activity increasing the
roughing rate of the remaining active particles. Just before the overall
activity drops only a fraction of the particles participates in the reaction
and the deactivation will happen very suddenly. Once these potentially few
particles eventually becomes sufficiently rough they will reduce and lose
activity hence completing the oscillation cycle. It should be noted that our
group has recently measured turnover frequencies of CO oxidation on Pt in
excess of $10^{6}$\,s$^{-1}$ per site corresponding to as little as $\sim$300 Pt
nanoparticles contributing to the reactivity just before the end of the high-
activity cycle.

\section{Conclusions}
Self-sustained CO oxidation oscillations at atmospheric pressure on Pt
nanoparticles ranging from from 3\,nm to 9\,nm have been measured with
high-quality mass-spectrometry. Full conversion was achieved on 0.1\,\%
geometrical Pt coverage and it is found likely that full conversion can be obtained on
extremely small amounts of Pt. It was found that the samples increased the
oscillation period over time and showed a strong temperature dependence. A
model is proposed where the nanoparticles go through repetitive
oxidation/reduction cycles which change the change rate of CO oxidation on the
particles.

\section{Experimental Methods}
\subsection{The $\mu$-reactor platform}
The experiments were performed in Si-based 20x16\,mm
$\mu$-reactors\cite{Henriksen2009}. The reactor consists of two inlets, a
mixing zone allowing for diffusional mixing of reactants, an outlet, a reactor
volume of $240\,$nL and a $5.4\,\mu$m wide, 3\,$\mu$m high and 1500\,$\mu$m
long capillary used for sniffing gases from the reactor volume. The reactor is
sealed by anodic bonding of a pyrex lid to the Si reactor and is able to
operate in a pressure range of 0--2.5\,bar. The reaction gases are supplied to
the two inlets by flow controllers capable of controlling the gas flow from 0--
10 mL/min The capillary flow is fed into a QMS
for analysis while any surplus of gas from the inlets is passed directly
through to the outlet via a pressure controller allowing control of reactor
pressure. The design makes sure that all gases exposed to the catalyst under
investigation are measured by the QMS ensuring an extremely high sensitivity of
the system. \begin{figure} \includegraphics[width=8cm]{Ib-reactor.png}
\caption{Image of the 20x16\,mm microreactor showing the circular reactor area
and gas channel system} \label{fgr:reactor} \end{figure} Before any
measurements are performed the reactor is pumped by a turbopump to minimize
contaminants in the system. When an evacuated reactor is mounted the base
pressure of the mass spectrometer chamber is $\sim5\times10^{-9}\,$mbar and in
operation at $1\,$bar in the reactor volume the pressure is
$\sim5\times10^{-7}\,$mbar in the QMS chamber.

@@

\subsection{Pt nanoparticle deposition}
Pt nanoparticles were deposited in the reactor using a gas-aggregation source
(Mantis Deposition Ltd., Nanogen 50). Pt clusters were formed by gas-phase
condensation of Pt atoms produced by impinging argon ions on a 99.99\% Pt
target in a magnetron sputter source. After condensation the ionised fraction
(60-80\%) of the clusters is size-selected by a quadropole mass selection
filter according to their mass-to-charge ratio. Using this setup Pt
nanoparticles with diameters in the range of 2--16\,nm can be produced
\cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after
size-selection soft-landed in the reactor volume of the microreactor. The
coverage was kept at approximately 0.1\% geometric coverage determined by
measuring the current on the reactor during deposition. After deposition the
reactors were anodically cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid
sintering of the nanoparticles while bonding.
